Compositional equivalences based on Open pNets

### 2021-10-20 2016-Weak-FH-Bisim/WeakBisim.tex
TODOs

\subsection{Open Automata}
 Open automata (OA) are not composition structures but they are made of transitions that are dependent of the actions of the holes, and they can reason on a set of variables (potentially with only symbolic values). 
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]\label{def:OT}
	\label{def:OpenTransitions}
	An \emph{open transition} (OT) over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$, a set $V$ of variables, and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\openrule
	{	\beta_j^{j\in J'}, \Pred, \Post}
	{s \OTarrow {\alpha}s'}
	\end{mathpar}
	Where $J'\subseteq J$ is the set of holes involved in the transition; $s, s'\in\mathcal{S}$ are states of the automaton; and $\beta_j$
        is a transition of the hole $j$, with $\Sortop(\beta_j)\in\Sort_j$. $\alpha$  is the resulting action of this open transition.
%        We call \emph{source variables} of an OT the union of all variables
%        in the different terms
%        $\beta_j$ and $\alpha$, and the state variables of $s$.
        \Pred\ is a predicate, \Post\ is a set of 
	assignments that are effective after the open transition, they are
        represented as a substitution function: $({x_k\gets e_k})^{k\in K}$.
Predicates and expressions of an open transition can refer to the variables in $V$, and
        in the different terms
        $\beta_j$ and $\alpha$. More precisely:
\[ \vars(\Pred) \subseteq V\cup\vars(\alpha)\cup
			{\displaystyle \bigcup_{j\in J'}\!\!\vars(\beta_j)}\quad\land\quad
\forall k.\, x_k\in V \land \vars(e_k)\subseteq V\cup\vars(\alpha)\cup
			{\displaystyle \bigcup_{j\in J'}\!\!\vars(\beta_j)}\]
The assignments are always applied simultaneously because the variables in $V$ can be in both sides ($x_k$s are distinct).
\TODO{$\forall k.\, x_k\in V \land \vars(e_k)$ pas clair!? c'est plutot cap?}
 Open transitions are identified
        modulo logical equivalence on their predicate. 
\end{definition}

@@

For all predicate $\Pred$, for all partial functions $\Post$ such that $V\cap \dom(\Post)=\emptyset$. We have: 
		 \begin{mathpar}
    \openrule
         {
           \set{\beta}, \Pred\,',\Post\,'}
          {t \OTarrow {\alpha} t'}\in\mathcal{T}
\implies
    \openrule
         {
           \set{\beta}\subst{\Post}, \Pred\,'\subst{\Post}\land\Pred,\Post\shortodot\Post\,'}
         {\raisemath{-2pt}{t \OTarrow {\alpha\subst{\Post}} {t'}}{}}
 \in\mathcal{T}
\end{mathpar}
\end{definition}

@@

Though the definition is simple, the fact that transitions are complex structures relating events must not be underestimated in order to understand the rest of the article. The first element of theory for open automata, i.e. the definition of a strong bisimulation, is given below.

@@

\end{mathpar}
 there exists an indexed set of  open transitions $OT_x^{x\in X} \subseteq \mathcal{T}_2$:
 \begin{mathpar}
%    \left( fresh \ \set{\alpha_i}, \set{b_j}, v_x.\ \
    \openrule
         {
           \beta_{j x}^{j\in J_{x}}, \Pred_{OT_x},\Post_{OT_x}}
         {t \OTarrow {\alpha_x} t_x}
%         \right)
\end{mathpar}
 such that  $\forall x.\, J'=J_{x}$ and there exists $\Pred_{s',t_x}$ such that $(s',t_x|\Pred_{s',t_x})\in 
 \mathcal{R}$
 and  \\
 $\Pred_{s,t} \land \Pred_{OT}\implies$\\
%\hspace{1cm}
 $\displaystyle{\bigvee_{x\in X}
   \left( \forall j. \beta_j=\beta_{jx}  \land \Pred_{OT_x}
     \land \alpha\!=\!\alpha_x \land  
     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\Post_{OT_x}}\right)}$
     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
     %     \right)$.
%\bigskip
%
% $\Pred_{s,t} \land \Pred_{OT}\implies\bigvee_{x\in X} \Pred_x$
%and\\
%%\hspace{1cm}
% $ \forall{x\in X}. \Pred_x \land \Pred_{s,t} \land \Pred_{OT} \Rightarrow
%   \left( \forall j. \beta_j=\beta_{jx}  \land \Pred_{OT_x}
%     \land \alpha\!=\!\alpha_x \land  
%     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\Post_{OT_x}}\right)$
%     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
%     %     \right)$.

@@

\label{Appendix:FullExample}
\TODO{Ludo: spec implem dans l annexe}

>>>>

\subsection{Open Automata}
 Open automata (OA) are not composition structures but they are made of transitions that are dependent of the actions of the holes, and they can reason on a set of variables (potentially with only symbolic values). 
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]\label{def:OT}
	\label{def:OpenTransitions}
	An \emph{open transition} (OT) over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$, a set $V$ of variables, and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\openrule
	{	\beta_j^{j\in J'}, \Pred, \Post}
	{s \OTarrow {\alpha}s'}
	\end{mathpar}
	Where $J'\subseteq J$ is the set of holes involved in the transition; $s, s'\in\mathcal{S}$ are states of the automaton; and $\beta_j$
        is a transition of the hole $j$, with $\Sortop(\beta_j)\in\Sort_j$. $\alpha$  is the resulting action of this open transition.
%        We call \emph{source variables} of an OT the union of all variables
%        in the different terms
%        $\beta_j$ and $\alpha$, and the state variables of $s$.
        \Pred\ is a predicate, \Post\ is a set of 
	assignments that are effective after the open transition, they are
        represented as a substitution function: $({x_k\gets e_k})^{k\in K}$.
Predicates and expressions of an open transition can refer to the variables in $V$, and
        in the different terms
        $\beta_j$ and $\alpha$. More precisely:
\begin{multline*}
 \vars(\Pred) \subseteq V\cup\vars(\alpha)\cup
			{\displaystyle \bigcup_{j\in J'}\!\!\vars(\beta_j)}\quad\land\\
\forall k.\, x_k\in V \quad\land\quad \forall k.\,\vars(e_k)\subseteq V\cup\vars(\alpha)\cup
			{\displaystyle \bigcup_{j\in J'}\!\!\vars(\beta_j)}
\end{multline*}
The assignments are applied simultaneously because the variables in $V$ can be in both sides ($x_k$s are distinct).
 Open transitions are identified
        modulo logical equivalence on their predicate. 
\end{definition}

@@

For all predicate $\Pred$ for all partial function $\Post$,if $V\!\cap \!\dom(\Post)=\emptyset$, we have: 
		 \begin{mathpar}
    \openrule
         {
           \set{\beta}, \Pred\,',\Post\,'}
          {t \OTarrow {\alpha} t'}\in\mathcal{T}
\implies
    \openrule
         {
           \set{\beta}\subst{\Post}, \Pred\,'\subst{\Post}\land\Pred,\Post\shortodot\Post\,'}
         {\raisemath{-2pt}{t \OTarrow {\alpha\subst{\Post}} {t'}}{}}
 \in\mathcal{T}
\end{mathpar}
\end{definition}

@@

Though the definition is simple, the fact that transitions are complex structures relating events must not be underestimated. The first element of theory for open automata, i.e. the definition of a strong bisimulation, is given below.

@@

\end{mathpar}
 there exists an indexed set of  open transitions $OT_x^{x\in X} \subseteq \mathcal{T}_2$:
 \begin{mathpar}
%    \left( fresh \ \set{\alpha_i}, \set{b_j}, v_x.\ \
    \openrule
         {
           \beta_{j x}^{j\in J_{x}}, \Pred_{OT_x},\Post_{OT_x}}
         {t \OTarrow {\alpha_x} t_x}
%         \right)
\end{mathpar}
 such that  $\forall x.\, J'=J_{x}$ and there exists $\Pred_{s',t_x}$ such that $(s',t_x|\Pred_{s',t_x})\in 
 \mathcal{R}$
 and  
\begin{multline*}
 \Pred_{s,t} \land \Pred_{OT}\implies\\
%\hspace{1cm}
 \displaystyle{\bigvee_{x\in X}
   \left( \forall j. \beta_j=\beta_{jx}  \land \Pred_{OT_x}
     \land \alpha\!=\!\alpha_x \land  
     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\Post_{OT_x}}\right)}
\end{multline*}
     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
     %     \right)$.
%\bigskip
%
% $\Pred_{s,t} \land \Pred_{OT}\implies\bigvee_{x\in X} \Pred_x$
%and\\
%%\hspace{1cm}
% $ \forall{x\in X}. \Pred_x \land \Pred_{s,t} \land \Pred_{OT} \Rightarrow
%   \left( \forall j. \beta_j=\beta_{jx}  \land \Pred_{OT_x}
%     \land \alpha\!=\!\alpha_x \land  
%     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\Post_{OT_x}}\right)$
%     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
%     %     \right)$.

@@

\label{Appendix:FullExample}
